The Spine of American Law: Digital History Methods and U.S. Legal Practice

### 2016-07-28 article/Funk-Mullen.Spine-of-Legal-Practice.Rmd
Response to Jason's comments

Curating a corpus to answer a specific question is an important difference from the way that digital history often describes its work. Speaking in broad terms, digital historical work---like all historical work---can begin either with a set of sources or with a question. Computational digital history is often conceived of as beginning with sources, particularly with large datasets such as the Google Books or Hathi Trust corpora. These large corpora are sometimes called "big data"---though, it must be emphasized, almost never by digital historians who actually work with them---on which "distant reading" can be practiced. While it is salutary for historians to have their research questions shaped by the broadest possible contexts, it is not apparent that digital historians can readily move from these omnibus corpora to answering the specific research questions that animate various historical fields. In this article we demonstrate an alternative approach, which we might unimaginatively label "medium data." The amount of legislation that governed American civil practice is impressive, since every state amended and re-enacted procedure statutes nearly every decade. But while a corpus of procedural legislation requires some computational sophistication, the techniques are far less complex than those derived for truly big data. We have gathered a large but narrowly constrained corpus centered on solving a well-defined research question. This corpus is large enough that digital historical methods provide results that a scholar could not obtain through traditional methods, but sufficiently circumscribed so as to directly address a discipline- and field-specific question.[^31]

>>>>

Curating a corpus to answer a specific question is one of two ways in which digital history can proceed. To quote Jason Heppler, digital history--like all historical work---can begin either "with a corpus looking for a question, or a question looking for a corpus." Computational text analysis in digital history is often conceived of as beginning with sources, particularly with large datasets such as the Google Books or Hathi Trust corpora. These large corpora are sometimes called "big data"---though it must be emphasized almost never by digital historians who actually work with them---on which "distant reading" can be practiced. While it is salutary for historians to have their research questions shaped by the broadest possible contexts, it is not apparent that digital historians can readily move from these omnibus corpora to answering the specific research questions that animate various historical fields. In this article we demonstrate an alternative approach, which we might unimaginatively label "medium data." The amount of legislation that governed American civil practice is impressive, since every state amended and re-enacted procedure statutes nearly every decade. But while a corpus of procedural legislation requires some computational sophistication, the techniques are far less complex than those derived for truly big data. We have gathered a large but narrowly constrained corpus centered on solving a well-defined research question.[^310] This corpus is large enough that digital historical methods provide results that a scholar could not obtain through traditional methods, but sufficiently circumscribed so as to directly address a discipline- and field-specific question.[^31]

[^310]: Our approach draws on an earlier generation of digital history which collected sources, as exemplified in *The Valley of the Shadow* project: William G. Thomas III and Edward L. Ayers, "The Differences Slavery Made: A Close Analysis of Two American Communities," *The American Historical Review* 108, no. 5 (December 1, 2003): 1299–1307, doi:10.1086/587017. See also Dan Cohen and Roy Rosenzweig, *Digital History: A Guide to Gathering, Preserving, and Presenting the Past on the Web* (Philadelphia: University of Pennsylvania Press, 2005), ch. 6, digital edition hosted at Roy Rosenzweig Center for History and New Media, George Mason University: <http://chnm.gmu.edu/digitalhistory/>.
